### 2020-09-17 paper/perfectGraphs.tex
Fixes

\subsubsection{Basic classes of perfect graphs}
Bipartite graphs are perfect, since we can color them with two colors. From the theorem of König we get that line graphs of bipartite graphs are also perfect \cite{Knig1916, GC03}. From the perfect graph theorem (\Cref{thm:pgt}) it follows that complements of bipartite graphs and complement of line graphs of bipartite graphs are also perfect. We will call these four classes \emph{basic}.

\subsubsection{2-join, Homogeneous Pair and Skew Partition}
A graph $G$ has a \emph{2-join} if its vertices can be partitioned into sets $V_1$, $V_2$, each of size at least three, and there are nonempty disjoint subsets $A_1, B_1 \subseteq V_1$ and $A_2, B_2 \subseteq V_2$, such that all vertices of $A_1$ are adjacent to all vertices of $A_2$, all vertices of $B_1$ are adjacent to all vertices of $B_2$, and these are the only edges between $V_1$ and $V_2$. When a graph $G$ has a 2-join, it can be decomposed onto two smaller graphs $G_1$, $G_2$, so that $G$ is perfect if and only if $G_1$ and $G_2$ are perfect \cite{Cornujols1985}.

@@

The proof of \Cref{thm:spgt} is a consequence of the \emph{Decomposition theorem}:
\begin{theorem}[Decomposition theorem, stated as Theorem 4.2 of \cite{GC03}]
	\label{thm:decomposition}
	Every Berge graph $G$ is basic or has a skew partition or a homogeneous pair, or either $G$ or $\overline{G}$ has a 2-join.
\end{theorem}

@@

The following is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but we will aim to show the intuition behind the algorithm.

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole or for shortest odd antihole (\Cref{AmenableHoles}).

@@

\begin{alg}[Test if $G$ contains a Pyramid]
	\label{alg:testPyramid}
	Input: A graph $G$.

@@

					\mIf{$s_1 = b_1$ \label{line:ShortestPathsStart}} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner \label{line:ShortestPathsEnd}
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices \lsx} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

Now we will proove the time complexity.

@@

Most obvious way to find a jewel is to enumerate all (possibly chordal) cycles of length 5 as $v_1, \ldots v_5$, check if it has all required nonedges and if it does, try to find a path $P$ as required. This gives us a time of $O(|V|^7)$. We could speed it up to $O(|V|^6)$ with more careful algorithm, but since whole Berge recognition algorithm takes time $O(|V|^9)$ and our testing showed that time it takes to test for jewels is negligible we decided against it.

@@

\subsection{Amenable holes.}
\label{AmenableHoles}
As we will see (\Cref{thm:amenableHoles}), if graphs $G$ and $\overline{G}$ contain no jewels, pyramids or configurations of type $\T_1, \T_2$ or $\T_3$, then every shortest odd hole in $G$ or $\overline{G}$ will have a special structure and we will call it amenable. This fact will help us determine if $G$ or $\overline{G}$ contain an odd hole. But first, let us introduce a few new definitions.

@@

Let us notice, that if $X$ is a cleaner for $C$, then $C$ is a clean hole in $G \setminus X$.

@@

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $\{v_8, v_9, v_1, v_2, v_3\}$ is an axample of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

\begin{theorem}[Theorem 8.1 of \cite{MC05}]
	\label{thm:amenableHoles}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest odd hole in $G$ is amenable.
	% \TODO{Any ideas on what else to say here? List all 9 steps?}
	% \TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

@@

\begin{alg}[List possible near cleaners]
	\label{alg:listNearCleaners}
	Input: A graph $G$.

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	\label{thm:92}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

\begin{alg}[Test possible near cleaner]
	\label{alg:testNearCleaner}
	Input: A graph $G$ containing no simple forbidden structure, and a subset $X \subseteq V(G)$.

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$ \label{line:definey2}
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2)$ \AND
						\lsx $r(x_2, y_1) \leq min(r(x_3, y_1), r(x_3, y_2))$ \label{line:manyconditions}}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	\label{thm:testNCWorks}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole, else there is no shortest odd hole $C$ in $G$, such that $X$ is a near-cleaner for $C$.
\end{theorem}
\begin{proof}
	Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The rest of the proof is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

@@

	By \Cref{thm:92} in line \ref{line:getR} we get a set $\mathcal{R}$, such that if $G$ contains an amenable hole $C$, one of the sets $X \in \mathcal{R}$ is a near-cleaner for $C$. In lines \ref{line:testXStart}-\ref{line:testXEnd} we test each $X \in \mathcal{R}$ and if an odd hole is reported, then by \Cref{thm:testNCWorks} $G$ is not Berge. If for all $X \in \mathcal{R}$ an odd hole is not reported, then by \Cref{thm:amenableHoles}, \Cref{thm:92} and \Cref{thm:testNCWorks} there is no odd hole in $G$. We run the same checks on $\overline{G}$ and if again we do not detect an odd hole, then $G$ is Berge and by \Cref{thm:spgt} $G$ is perfect.

>>>>

\subsubsection{Basic classes of perfect graphs}
Bipartite graphs are perfect, since we can color them with two colors. From the theorem of König \cite{Knig1916} we get that line graphs of bipartite graphs are also perfect \cite{GC03}. From the perfect graph theorem (\Cref{thm:pgt}) it follows that complements of bipartite graphs and complement of line graphs of bipartite graphs are also perfect. We will call these four classes \emph{basic}.

\subsubsection{2-join, Homogeneous Pair and Skew Partition}
A graph $G$ has a \emph{2-join} if and only if its vertices can be partitioned into sets $V_1$, $V_2$, each of size at least three, and there are nonempty disjoint subsets $A_1, B_1 \subseteq V_1$ and $A_2, B_2 \subseteq V_2$, such that all vertices of $A_1$ are adjacent to all vertices of $A_2$, all vertices of $B_1$ are adjacent to all vertices of $B_2$, and these are the only edges between $V_1$ and $V_2$. When a graph $G$ has a 2-join, it can be decomposed onto two smaller graphs $G_1$, $G_2$, so that $G$ is perfect if and only if $G_1$ and $G_2$ are perfect \cite{Cornujols1985}.

@@

The proof of \Cref{thm:spgt} is a consequence of the \emph{Decomposition theorem}:

\begin{theorem}[Decomposition theorem, stated as Theorem 4.2 of \cite{GC03}]
	\label{thm:decomposition}
	Every Berge graph $G$ is basic or has a skew partition or a homogeneous pair, or either $G$ or $\overline{G}$ has a 2-join.
\end{theorem}

@@

The following section is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but we will aim to show the intuition behind the algorithm.

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole or for shortest odd antihole (\Cref{AmenableHoles}). If we find no such near-cleaners, then, as we will see, the graph is Berge.

@@

\begin{alg}[Test if $G$ contains a pyramid]
	\label{alg:testPyramid}
	Input: A graph $G$.

@@

					\mIf{$s_1 = b_1$ \label{line:ShortestPathsStart}} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1$-$S_1(m)$-$m$-$T_1(m)$-$b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner \label{line:ShortestPathsEnd}
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\ls color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices \lsx} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j) \in good\_pairs_{i, 2}$}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

Now we will prove the time complexity.

@@

Most obvious way to find a jewel is to enumerate all (possibly chordal) cycles of length 5 as $v_1, \ldots v_5$, check if it has all required nonedges and if it does, try to find a path $P$ as required. This gives us a time of $O(|V|^7)$. We could speed it up to $O(|V|^6)$ with more careful algorithm (see Algorithm 3.1 of \cite{MC05}), but since whole Berge recognition algorithm takes time $O(|V|^9)$ and our testing showed that time it takes to test for jewels is negligible we decided against it.

@@

When graphs $G$ and $\overline{G}$ contain no pyramids, no jewels and no configurations of type $\T_1, \T_2$ or $\T_3$, we will say that graph $G$ contains no \emph{simple forbidden structures}.

\subsection{Amenable holes.}
\label{AmenableHoles}
As we will see (\Cref{thm:amenableHoles}), if graph $G$ contains no simple forbidden structures, then every shortest odd hole in $G$ or $\overline{G}$ will have a special structure and we will call it amenable. This fact will help us determine if $G$ or $\overline{G}$ contain an odd hole. But first, let us introduce a few new definitions.

@@

\noindent Let us notice, that if $X$ is a cleaner for $C$, then $C$ is a clean hole in $G \setminus X$.

@@

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $\{v_8, v_9, v_1, v_2, v_3\}$ is an example of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

\begin{theorem}[Theorem 8.1 of \cite{MC05}]
	\label{thm:amenableHoles}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no pyramids, no jewels and no configurations of types $\T_1, \T_2$ or $\T_3$. Then every shortest odd hole in $G$ is amenable.
	% \TODO{Any ideas on what else to say here? List all 9 steps?}
	% \TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

@@

\begin{alg}[List possible near cleaners]
	\label{alg:listNearCleaners}
	Input: A graph $G$ containing no simple forbidden structures.

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	\label{thm:92}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no pyramids, no jewels and no configurations of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

\begin{alg}[Test possible near cleaner]
	\label{alg:testNearCleaner}
	Input: A graph $G$ containing no simple forbidden structures, and a subset $X \subseteq V(G)$.

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1$-$x_2$-$x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$ \label{line:definey2}
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2)$ \AND
						\lsx $r(x_2, y_1) \leq min(r(x_3, y_1), r(x_3, y_2))$ \label{line:manyconditions}}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	\label{thm:testNCWorks}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no pyramids, no jewels and no configurations of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole, else there is no shortest odd hole $C$ in $G$, such that $X$ is a near-cleaner for $C$.
\end{theorem}
\begin{proof}
	Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The rest of the proof is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

@@

	By \Cref{thm:92} in line \ref{line:getR} we get a set $\mathcal{R}$, such that if $G$ contains an amenable hole $C$, one of the sets $X \in \mathcal{R}$ is a near-cleaner for $C$. In lines \ref{line:testXStart}-\ref{line:testXEnd} we test each $X \in \mathcal{R}$ and if an odd hole is reported, then by \Cref{thm:testNCWorks} $G$ is not Berge. If for all $X \in \mathcal{R}$ no odd holes are reported, then by \Cref{thm:amenableHoles}, \Cref{thm:92} and \Cref{thm:testNCWorks} there are no odd holes in $G$. We run the same checks on $\overline{G}$ and if again we do not detect an odd hole, then $G$ is Berge and by \Cref{thm:spgt} $G$ is perfect.
